---
title: "Mental Health Illnesses among Adults with And without A History of Cancer in the United States"
subtitle: "GLHLTH 562, Spring 2023"
author: "Meng Chen, Shurui Wang"
institute: "Duke Global Health Institute"
date: "2023/04/10"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css,custom_deck.css]
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
    seal: false

---

class: titleSlide 
class: center, middle
```{r setup, include=FALSE}
library(tidyverse)
library(palmerpenguins)
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.width=9, fig.height=3.5, fig.retina=3, out.width = "100%", cache = FALSE, echo = TRUE, messAge = FALSE, warning = FALSE, hiline = TRUE)

library(tidyverse)
library(biostat3)
library(broom)
library(ggplot2)
library(xaringan)
library(xaringanthemer)
library(xaringanExtra)
library(patchwork)
```

```{r r xaringan-themer, include=FALSE, warning=FALSE}
style_duo_accent(
  primary_color = "#1889c9",
  secondary_color = "#fec52e",
  header_font_google = google_font("Source Sans Pro"),
  text_font_google   = google_font("Libre Franklin", "300", "300i"),
  code_font_google   = google_font("Anonymous Pro"),
  base_font_size = "30px",
  text_font_size = "1rem",
  header_h1_font_size = "2.5rem",
  header_h2_font_size = "2rem",
  header_h3_font_size = "1.25rem",
  padding = "8px 32px 8px 32px",
)
```

## `r rmarkdown::metadata$title`

#### `r rmarkdown::metadata$subtitle`

#### `r rmarkdown::metadata$author`

#### `r rmarkdown::metadata$date`


---
class: left

## Introduction

```{r xaringan-panelset1, echo=FALSE}
xaringanExtra::use_panelset()
```

.panelset[

.panel[.panel-name[Background]

 - **Over 18 million** cancer survivors and **over 1.9 million** new cancer cases in 2023. 
 - Higher risks of mental health illness for individuals with a history of cancer.
 - Studies mainly focused on survivors of **specific types** of cancer or were conducted in **other countries**.

]

.panel[.panel-name[Aim]

 1. To examine the association between any mental health illness (AMI) and a history cancer in the US.
 2. To examine the risk factors of AMI among individuals with a history of cancer. 

]

]


---
class: left

## Methods

 - **Database**: 
  - National Survey on Drug Use and Health (NSDUH)-- open-access; annual; nationwide; mental health 
 - **Variable**: 
  - Outcome: Any mental illness (AMI)
  - Exposure: Cancer, Drug, Mental health service (MH_service), Age, Sex, Education, Race
 - **Analysis**: 
  - Baseline characteristics of the whole - barplot
  - Difference between non_cancer and cancer - barplot
  - Association between AMI and cancer status - regression

---

class: left

## Results

```{r wraggling1,echo=FALSE, warning=FALSE, message=FALSE}
NSDUH <- read.csv(here::here("data", "NSDUH_final_data.csv"))


variables <- c("Cancer", "Drug", "MH_service", "Age", "Sex", "Education", "Race")

# Create a function to map subtype values to labels
map_subtype_labels <- function(variable_name, subtype_value) {
  labels <- list(
    Cancer = c("0" = "Non_cancer", "1" = "Cancer"),
    Drug = c("0" = "No illicit drug use", "1" = "Illicit drug use"), 
    MH_service= c("1" = "Inpatient", "2" = "Outpatient", "3"="Medication", "4"="No mental health services"),
    Age = c("1" = "18-34 yrs", "2" = "25-49 yrs", "3"="50-64 yrs", "4"="≥65 yrs"),
    Education= c("1" = "Less than high school", "2"= "High school graduate", "3"="Some college", "4"="College graduate or higher"),
    Sex = c("1" = "Male", "2" = "Female"),
    Race = c("1" = "Non-Hispanic white", "2"="Non-Hispanic black", "3"="Non-Hispanic other", "4"="Hispanic")
  )
  return(labels[[variable_name]][[subtype_value]])
}


# Create a function to calculate the percentages for each variable
calculate_percentages <- function(variable_name) {
  data_cohort <- NSDUH %>%
    count(!!sym(variable_name)) %>%
    mutate(Percentage = n / sum(n) * 100) %>%
    rename(Subtype = !!sym(variable_name)) %>%
    mutate(Category = variable_name)
  return(data_cohort)
}

plot_NSDUH <- bind_rows(lapply(variables, calculate_percentages))

# Add the Label column
plot_NSDUH$Label <- mapply(map_subtype_labels, plot_NSDUH$Category, as.character(plot_NSDUH$Subtype))

# Create a stacked bar plot
ggplot(plot_NSDUH, aes(x = Category, y = Percentage, fill = Label)) +
  geom_bar(stat = "identity") +
  labs(title = "Baseline Characteristics of the Whole Cohort", x = NULL, y = "Percentage") +
  scale_fill_viridis_d(guide = guide_legend(title = "Subtype"), option = "viridis") +
  theme_minimal() +
  coord_flip()


```

---

class: left

## Results

```{r AMI_barplot, echo=FALSE}

# calculate percentages for each group
percentages <- aggregate(AMI ~ Cancer, data = NSDUH, FUN = function(x) round(mean(x) * 100, 1))

# plot the bar plot
ggplot(percentages, aes(x = factor(Cancer), y = AMI)) +
  geom_bar(stat = "identity", fill = "#7AA9DD", alpha = 0.7) +
  ylim(0, 30) +
  geom_text(aes(label = paste0(AMI, "%")), vjust = -0.5) +
  labs(x = " ", y = " ") +
  ggtitle("Percentage of Having Any Mental Illness among Individuals with and without Cancer") +
  theme_minimal()+
  scale_x_discrete(labels = c("Non_cancer", "Cancer"))
```
---
class: left

## Results
```{r OR-plot, echo=FALSE}
NSDUH_count_all <- NSDUH %>%
  gather(variable, value, -Cancer, -AMI) %>%
  count(variable, value, Cancer, AMI) %>%
  group_by(variable, value, Cancer) %>%
  mutate(Percentage = n / sum(n) * 100) %>%
  ungroup()

# Define a function to create a bar plot for each exposure variable
create_bar_plot <- function(variable_name, index) {
  custom_colors <- c("#7AA9DD", "#2C3E50")
  plot <- ggplot(NSDUH_count_all %>% filter(variable == variable_name),
                 aes(x = value, y = Percentage, fill = as.factor(AMI))) +
    geom_bar(stat = "identity", position = "dodge") +
    facet_wrap(~ Cancer, labeller = as_labeller(c("0" = "Non_cancer", "1" = "Cancer"))) +
    scale_fill_manual(values = custom_colors)+
    labs(title = paste(index),
         x = variable_name,
         y = NULL,
         fill = "AMI") +
    theme_minimal() +
    theme(legend.position = "right")
  return(plot)
}

# Create a list of plots
exposure_variables <- c("Drug", "MH_service", "Age", "Sex", "Education", "Race")
bar_plots <- lapply(seq_along(exposure_variables), function(i) create_bar_plot(exposure_variables[i], i))


# Combine the plots with patchwork
combined_plot <- (bar_plots[[1]] + bar_plots[[2]] + bar_plots[[3]] + bar_plots[[4]] + bar_plots[[5]] + bar_plots[[6]]) +
  plot_annotation(title = "Percentage of People with AMI between People with and without Cancer Based on Different Exposure Status",
                  theme = theme(plot.title = element_text(hjust = 0.5)))

# Display the combined plot
print(combined_plot)


```


---
class: left

## Results

```{r wrangling_whole, include=FALSE}
NSDUH_whole <- NSDUH %>%
  mutate(Drug = factor(Drug))%>%
  mutate(MH_service = factor(MH_service))%>%
  mutate(Age = factor(Age))%>%
  mutate(Sex = factor(Sex))%>%
  mutate(Education = factor(Education))%>%
  mutate(Race = factor(Race))%>%
  mutate(AMI = factor(AMI))%>%
  mutate(Cancer = factor(Cancer))


# Run the model
whole_logit <- glm(AMI ~ Cancer + Drug + MH_service + Age + Sex + Education + Race , data = NSDUH_whole, family = binomial(link = "logit"))

whole_logit_OR<- whole_logit %>%
  tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)

whole_logit_OR
```


<div style="height: 500px; overflow: scroll;">
```{r fp_whole, echo=FALSE,include=FALSE}
library(ggplot2)
library(forestplot)
library(tidyverse)
library(readxl)

OR_reference <- read_excel(here::here("data", "fp.xlsx"))


row_name1 <- cbind(c("Predictors", OR_reference$Predictors), c("Odds Ratios (95% CI)", OR_reference$`OR (95%CI)`), c("P value", OR_reference$`P value`))

OR <- rbind(rep(NA,3),OR_reference)
```

```{r fp2_whole,fig.height=6,echo=FALSE}

forestplot(labeltext=row_name1,
           OR$OR,OR$`CL-D`,OR$`CL-U`,
           zero=1,
           xlog=F,
           is.summary = c(T,T,F,F,T,F,F,T,F,F,F,F,T,F,F,T,F,F,F,F,T,F,F,F,F,T,F,F,F,F),
           graph.pos=4, # 森林图出现在表格第几列
           boxsize = 0.3,  # 方块大小
           fnci_norm = fpDrawCircleCI, # 中间图形形状
           line.margin = unit(5, "mm"), # 线边距
           lineheight = unit(18, "mm"), # 设置行距
           colgap = unit(7,"mm"), # 列间距
           col = fpColors(box = "grey0", lines = "grey0", summary = "grey0"),
           xticks = c(0,1,2), # x轴坐标
           clip = c(0.2, 2.3),  # 置信区间范围， 超过会变成箭头
           lwd.zero =2, # 参考线粗细
           lwd.xaxis =1,
           lty.ci = 7, # 误差条的线性
           lwd.ci = 1, #误差条的线宽度
           ci.vertices.height = 0.15,
           gp = gpar(fontsize = 8))%>% # 误差条末端长度
  fp_add_lines() %>%
  fp_set_style(box = "royalblue",
               line = "darkblue",
               summary = "royalblue",
               align = "lrrr",
               hrz_lines = "#999999") %>%
  fp_set_zebra_style("#EFEFEF")

```
</div>

---
class: left

## Results

```{r wrangling_cancer, include=FALSE}
# Subset the data to include only cases with Cancer
Cancer <- subset(NSDUH, Cancer == 1) %>%
  mutate(Drug = factor(Drug))%>%
  mutate(MH_service = factor(MH_service))%>%
  mutate(Age = factor(Age))%>%
  mutate(Sex = factor(Sex))%>%
  mutate(Education = factor(Education))%>%
  mutate(Race = factor(Race))%>%
  mutate(AMI = factor(AMI))


# Run the model
Cancer_logit <- glm(AMI ~ Drug + MH_service + Age + Sex + Education + Race , data = Cancer, family = binomial(link = "logit"))

Cancer_logit_OR<- Cancer_logit %>%
  tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)

Cancer_logit_OR

```

<div style="height: 500px; overflow: scroll;">
```{r fp_cancer, echo=FALSE,include=FALSE}

OR_cancer <- read_excel(here::here("data", "fp_cancer.xlsx"))


row_name2 <- cbind(c("Predictors", OR_cancer$Predictors), c("Odds Ratios (95% CI)", OR_cancer$`OR (95%CI)`), c("P value", OR_cancer$`P value`))

OR_cancer <- rbind(rep(NA,3),OR_cancer)
```

```{r fp2_cancer,fig.height=6,echo=FALSE}

forestplot(labeltext=row_name2,
           OR_cancer$OR,OR_cancer$`CL-D`,OR_cancer$`CL-U`,
           zero=1,
           xlog=F,
           is.summary = c(T,T,F,F,T,F,F,F,F,T,F,F,T,F,F,F,F,T,F,F,F,F,T,F,F,F,F),
           graph.pos=4, # 森林图出现在表格第几列
           boxsize = 0.3,  # 方块大小
           fnci_norm = fpDrawCircleCI, # 中间图形形状
           line.margin = unit(5, "mm"), # 线边距
           lineheight = unit(18, "mm"), # 设置行距
           colgap = unit(7,"mm"), # 列间距
           col = fpColors(box = "grey0", lines = "grey0", summary = "grey0"),
           xticks = c(0,1,2), # x轴坐标
           clip = c(0.2, 2.3),  # 置信区间范围， 超过会变成箭头
           lwd.zero =2, # 参考线粗细
           lwd.xaxis =1,
           lty.ci = 7, # 误差条的线性
           lwd.ci = 1, #误差条的线宽度
           ci.vertices.height = 0.15,
           gp = gpar(fontsize = 8))%>% # 误差条末端长度
  fp_add_lines() %>%
  fp_set_style(box = "royalblue",
               line = "darkblue",
               summary = "royalblue",
               align = "lrrr",
               hrz_lines = "#999999") %>%
  fp_set_zebra_style("#EFEFEF")

```

</div>





